\documentclass{tufte-handout}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{chains}
% Define style for nodes
\tikzstyle{every node}=[circle, draw, fill=white,
                        inner sep=0pt, minimum width=9pt,font=\footnotesize\sf]
\tikzstyle{every picture}= [scale=.5,baseline=(current bounding box.center)]
\usetikzlibrary{automata}
\usetikzlibrary{patterns}
\usepackage{booktabs}
\usepackage{siunitx}
\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand\GITAuthorDate{no git info found}\newcommand\GITAbrHash{-}}  % For version control

\title{Heaps}
\author{}
\date{\GITAuthorDate, rev. \GITAbrHash}

\begin{document}
\maketitle

\section{\textbf{Description}}
This assignment is an exam question about understanding a given piece of code.

\section{\textbf{Deliverable}}
\begin{enumerate}
\item A PDF with your answers.
    Please add all the names and itu-email addresses of the group members as authors.
\end{enumerate}


\section{\textbf{Question 1}: from exam 180528 Question 2}
  \begin{figure}
    %\begin{tikzpicture}[remember picture,overlay] 
    %  \draw [fill=red!5]
    %  (current page.south west) rectangle (current page.north east); 
    %\end{tikzpicture}
    
    \begin{lstlisting}[basicstyle=\ttfamily,backgroundcolor=\color{white},
      frame=single,rulecolor=\color{gray!20},framesep=10pt,
      linewidth=12cm,numbers= left]
import edu.princeton.cs.algs4.*;

public class S
{
    public int[] a;
    int sz;

    public S(int cap)
    { a = new int[cap]; sz = 0;}

    public void push(int item)
    {
	if (sz == a.length) shift();
	a[sz] = item;
	sz++;
    }
    public int pop()
    {   
	sz--;
	return a[sz];
    }

    private void shift()
    {
	int[] tmp = new int[a.length];
	for (int i = 0; i<a.length - 1; i++) tmp[i] = a[i+1];
	a = tmp;
	sz--;
    }
}
    \end{lstlisting}
    \caption{\label{fig: S-java} Class {\tt S} (for Strange Stack), java version.}
  \end{figure}


  \begin{figure}
    %\begin{tikzpicture}[remember picture,overlay] 
    %  \draw [fill=red!5]
    %  (current page.south west) rectangle (current page.north east); 
    %\end{tikzpicture}
    
    \begin{lstlisting}[basicstyle=\ttfamily,backgroundcolor=\color{white},
      frame=single,rulecolor=\color{gray!20},framesep=10pt,
      linewidth=12cm,numbers= left]
class S:
    def __init__(self, capacity):
        self.a = [0] * capacity;
        self.sz = 0

    def push(self, item):
        if self.sz == len(self.a):
            self.shift()
        self.a[self.sz] = item
        self.sz += 1

    def pop(self):
        self.sz -= 1
        return self.a[self.sz]

    def shift(self):
        tmp = [0] * len(self.a)
        for i in range(0, len(self.a) - 1):
            tmp[i] = self.a[i+1]
        self.a = tmp
        self.sz -= 1
    \end{lstlisting}
    \caption{\label{fig: S-python} Class {\tt S} (for Strange Stack), python version.}
  \end{figure}
    
    The next few questions all concern the class defined in fig.~\ref{fig: S-java} and \ref{fig: S-python}.

    \subsection*{(a) 1Pt} Class $S$ is some kind of fixed-capacity stack-like data structure.
      When an object of class $S$ runs out of capacity, it
      \begin{enumerate}
      \item forgets the oldest element.\\
      \item ignores the next push.\\
      \item   resizes itself so as to make room for more elements. \\ 
      \item   return the smallest element. \\
 
      \end{enumerate}
    \subsection*{(b) 1Pt}
      What is the result of the following operations?
      \begin{minipage}{7cm}
      \begin{verbatim}
	// java 
	int N = 4;	
	S s = new S(N);
	for (int i = 0; i < N    ; i++) 
	    s.push(i);
	for (int i = 0; i < N    ; i++) 
	    StdOut.print(s.pop());
	for (int i = 0; i < N + 1; i++) 
	    s.push(i);
	for (int i = 0; i < N    ; i++) 
	    StdOut.print(s.pop());
      \end{verbatim}
      \end{minipage}
%
      \begin{minipage}{7cm}
      \begin{verbatim}
      # python3
      N = 4;
      s = S(N)
      for i in range(N):
          s.push(i)
      for i in range(N):
          stdio.write(s.pop())
      for i in range(N + 1): 
          s.push(i)
      for i in range(N): 
          stdio.write(s.pop())
      \end{verbatim}
      \end{minipage}
       
      (Ignore whitespace such as blanks or newlines.)
    \subsection*{(c) 1Pt}      
      Draw the data structure after the following operations:
      (This means ``\emph{at the end} of the operations,''
        not ``\emph{after each} operation,''  so you need to draw only a single picture.
        Make sure you draw the entire data structure.
        Make sure to include all instance variables.
        )
      \begin{minipage}{5cm}
      \begin{verbatim}
      // java
      S s = new S(3);
      s.push(5);
      s.push(6);
      s.pop();
      \end{verbatim}
      \end{minipage}
%
      \begin{minipage}{5cm}
      \begin{verbatim}
      # python 3
      s = S(3)
      s.push(5)
      s.push(6)
      s.pop()
      \end{verbatim}
      \end{minipage}

      \subsection*{(d) 1Pt}
      Assume I initialised an $S$-object with capacity $K$. 
      Now I perform a sequence of $N$ pushes.
      What is the worst-case (also the worst possible dependency of $K$ on $N$) running time of a single \texttt{push}?  (Choose the smallest correct estimate.)
      \begin{enumerate}
        \item $O(\log N)$. \\
        \item $O(N)$. \\
        \item $O(N\log N)$. \\
        \item  $O(1)$. \\
      \end{enumerate} 

      \subsection*{(e) 1Pt}
      Assume I initialised an $S$-object with capacity $K$. 
      Now I perform a sequence of $N$ pushes, followed by $N$ pops.
      What is the worst-case (also the worst possible dependency of $K$ on $N$) running time of a single \texttt{pop}?  (Choose the smallest correct estimate.)
      \begin{enumerate}
        \item $O(\log N)$. \\
        \item $O(N)$. \\
        \item $O(N\log N)$. \\
        \item  $O(1)$. \\
      \end{enumerate} 
      \subsection*{(f) 1Pt}
      Assume I initialised an $S$-object with capacity $K$. 
      Now I perform $N$ pushes and $N+1$ pops, where $N\leq K$.
      What happens? \textbf{The java and the python version behave differently, indicate what you answer to}
      \begin{enumerate}
	\item At least one of the pushes or pops takes linear time. \\
	\item The program aborts with a runtime error or exception. \\
	\item The underlying array resizes to $\frac{1}{4}$ of its size. \\
	\item The smallest element is returned. \\
      \end{enumerate} 

      \subsection*{(g) 1Pt}
     Add a method with the following signature that checks if the data structure is empty.

      \begin{minipage}{6cm}
      \begin{verbatim}
      // java
      public boolean isEmpty()
      \end{verbatim}
      \end{minipage}
%
      \begin{minipage}{6cm}
      \begin{verbatim}
      # python 3
      def is_empty(self):
      \end{verbatim}
      \end{minipage}

    \emph{Don't change any other methods in class {\tt S} and don't introduce new instance variables to {\tt S}.}
     
      \subsection*{(h) 1Pt}
      Extend class {\tt S} with the following signature that removes the element that was first pushed (i.e., the `oldest' element).
      The method returns nothing.
    \emph{Don't change any other methods in class {\tt S} and don't introduce new instance variables to {\tt S}.}

    \medskip
      \begin{minipage}{6cm}
      \begin{verbatim}
      // java
      public void remove_oldest()
      \end{verbatim}
      \end{minipage}
%
      \begin{minipage}{6cm}
      \begin{verbatim}
      # python 3
      def remove_oldest(self):
      \end{verbatim}
      \end{minipage}

      \subsection*{(i) 2Pt}
      Explain how to change the implementation of the data structure so that push and pop take constant time in the worst case.
      The functionality must be unchanged.
      If $K$ is the capacity, then the constructor should take $O(K)$ time and the data structure should take $O(K)$ space.

\end{document}